---
title: "Supply Dispatch Optimization: Test Report"
output: pdf_document
date: "`r format(Sys.time(),'%d-%m-%Y')`"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	error = F,
	fig.align = "center",
	fig.height = 5,
	fig.pos = "H",
	fig.width = 8.5,
	message = FALSE,
	warning = FALSE,
	dpi = 300,
	ft.left = 2.5
)
```


```{r}
# NOTE: RUN SHOULD BE AFTER RUNNING THE OPTIMIZATION!
```

# Basic "Sniff Test"

Just to know everything is ok. In **bold** there is the expected result:

No mine opened twice: **0 1**
```{r}
df_results %>%  group_by(d) %>% 
  reframe(x=sum(mine_opened)) %>% pull(x) %>% range()
```

Number of mines opened and total number of mines: **A reasonable number, above 0 and below 100**

```{r}
sum(df_results$mine_opened)
nrow(deposit)
```

Number of mines with production: **Should be same as mines opened**
```{r}
df_results %>% filter(tons_extracted>0.00001) %>% pull(d) %>% unique() %>% length()
```


Production at last period vs demand. Note there is existing capacity: **Equal or really close**

```{r}
sum(df_results$capacity_added)+sum(deposit$prod_rate)
demand$Demand[t_size]
```

Total extraction vs reserves: **Extraction (first) should be less than reserves**

```{r}
sum(df_results$tons_extracted1)+sum(df_results$tons_extracted2);
sum(deposit$reserve)
```

No production or extraction at unopened mines: **Should be 0**

```{r}
closed_d <- df_results %>% group_by(d) %>%
  reframe(x=sum(mine_opened)) %>% filter(x==0) %>% pull(d)
df_results %>% filter(d %in% closed_d) %>%
  reframe(tons_extracted=sum(tons_extracted),
          capacity_added=sum(capacity_added))
```

Max production rate achieved vs ramp up limit: **First should be lower/equal than second**

```{r}
df_results %>% reframe(x=max(cap_total)) %>% pull(x)
max(deposit$max_prod_rate) # bigM_extract
```

Test sequential (order) in extraction within each deposit, between shares (lower and higher cost). Extraction should always occurs first in the first half, and then the second part (more expensive).

```{r}
df_results %>% 
  filter(tons_extracted>0.001) %>% 
  left_join(deposit) %>% 
  mutate(first_half=case_when(
    total_extraction>=reserve-1 ~ "Depletion", # mnargin of 1 as they represent really close
    tons_extracted1>tons_extracted2 ~ "First half",
    T ~ "Second half")) %>%
  ggplot(aes(t,Deposit_Name))+
  geom_tile(aes(fill=first_half))+
    ggforce::facet_col(facets = vars(Resource_Type), 
                     scales = "free_y", 
                     space = "free") +
  scale_fill_manual(values = c("darkred","darkblue", "darkgreen"))+
  coord_cartesian(expand = F)+
  labs(x="",y="",fill="Extraction in")
```



## Shadow prices or dual model results

Should be exactly the **same/equal** as mixed integer results.


```{r}
# Objective value:
# compare MIP (gurobi) solution against LP (glpk)
# result$objective_value
# result_LP$objective_value+sum(cost_opening*openings)
```

Total mineral extracted: 

```{r}
sum(df_results$tons_extracted1)
sum(df_results_LP$tons_extracted1)
sum(df_results$tons_extracted2)
sum(df_results_LP$tons_extracted2)
```

Total capacity added: 

```{r}
sum(df_results$capacity_added)
sum(df_results_LP$capacity_added)
```

Mineral extraction detail:

```{r}
df_results %>% rename(value_orig=tons_extracted1) %>% 
  left_join(df_results_LP) %>% dplyr::select(d,t,value_orig,tons_extracted1) %>% 
  mutate(diff=abs(value_orig-tons_extracted1)) %>% arrange(desc(diff)) %>% head()
df_results %>% rename(value_orig=tons_extracted2) %>% 
  left_join(df_results_LP) %>% dplyr::select(d,t,value_orig,tons_extracted2) %>% 
  mutate(diff=abs(value_orig-tons_extracted2)) %>% arrange(desc(diff)) %>% head()
```

Capacity added detail:

```{r}
df_results %>% rename(value_orig=capacity_added) %>% 
  left_join(df_results_LP) %>% dplyr::select(d,t,value_orig,capacity_added) %>% 
  mutate(diff=abs(value_orig-capacity_added)) %>% arrange(desc(diff)) %>% head()
```


# Model Results

A brief glance at model results

```{r}
theme_set(theme_bw(16)+ theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(),axis.title.y=element_text(angle=0,margin=margin(r=0))))
```

## Production over time

```{r}
df_results %>% 
  left_join(deposit, by="d") %>% 
  filter(t<2051) %>% 
  group_by(Country,t) %>% 
  reframe(tons_extracted=sum(tons_extracted)) %>% ungroup() %>% 
  rename(d=Country) %>% 
  # aggregate big countries
  group_by(d) %>% mutate(total_prod=sum(tons_extracted)) %>% ungroup() %>% 
  group_by(t) %>% mutate(share_prod=total_prod/sum(total_prod)) %>% ungroup() %>% 
  mutate(d=if_else(share_prod>0.02,d,"Other countries")) %>% 
  group_by(d,t) %>% 
  reframe(tons_extracted=sum(tons_extracted)) %>% ungroup() %>% 
# end aggregate other countries
    # mutate(d=factor(d)) %>% 
  ggplot(aes(t,tons_extracted,fill=d,group=d))+
  geom_area()+
  coord_cartesian(expand = F)+
  labs(x="",y="Production [ktons]",fill="Country")+
  theme(
    # legend.position = "none",
        axis.text.x = element_text(hjust=1))
if (save_figures){ f.fig.save(sprintf(fig_name,"production"))}

# With Resource type now
df_results %>% 
  left_join(deposit, by="d") %>% 
  filter(t<2051) %>% 
  group_by(Resource_Type,t) %>% 
  reframe(tons_extracted=sum(tons_extracted)) %>% ungroup() %>% 
  rename(d=Resource_Type) %>% 
  # mutate(d=factor(d)) %>% 
  ggplot(aes(t,tons_extracted,fill=d,group=d))+
  geom_area()+
  coord_cartesian(expand = F)+
  scale_fill_manual(values=resource_colors)+
  labs(x="",y="Production [ktons]",fill="Resource \n Type")+
  theme(
    # legend.position = "none",
        axis.text.x = element_text(hjust=1))
if (save_figures){ f.fig.save(sprintf(fig_name,"productionType"))}

```

## Map cumulative production
```{r}
# World map with polygons
map1 <- map_data('world')

data_map <- df_results %>% 
  filter(t<2051) %>% group_by(d) %>% 
  reframe(tons_extracted=sum(tons_extracted)) %>% ungroup() %>% 
  filter(tons_extracted>0.001) %>%
  left_join(deposit, by="d") %>% 
  filter(!str_detect(Resource_Type,"Other"))
  
ggplot(data_map) +
  # base map
  theme_minimal() +
  geom_polygon(data = map1, 
               mapping = aes(x = long, 
                             y = lat, 
                             group = group),
               col = 'gray', fill="white") +
  coord_fixed(1.4, xlim = c(-140,160), ylim=c(-70,70))+
   scale_y_continuous(breaks = NULL,name = "")+
   scale_x_continuous(breaks = NULL,name = "")+
  theme(panel.grid = element_blank())+
  # data
  geom_point(aes(x = Longitude, y = Latitude,
                 size = tons_extracted,col=Resource_Type), alpha = 0.7) +
  scale_color_manual(values = resource_colors)+
  # scale_size_continuous(trans="log10") +
  labs(col="Resource \n Type",
       size = "Total tons extracted \n 2022-2050 [ktons]")
```



## Opened Mines over time

```{r}
df_results %>% 
  filter(t<2051) %>%
  rename(value=mine_opened) %>% 
  mutate(value=factor(value)) %>% 
  left_join(deposit,by="d") %>% 
  ggplot(aes(t,d,fill=value))+
  geom_tile()+
    ggforce::facet_col(facets = vars(Resource_Type), 
                     scales = "free_y", 
                     space = "free") +
  coord_cartesian(expand = F)+
  scale_y_continuous(breaks = NULL)+
  scale_fill_manual(values = c("0" = "white", "1" = "red"))+
  labs(x="",y="Deposit ",fill="",caption="Each row is a deposit. Red indicates opening.")+
  theme(legend.position = "none")
if (save_figures){f.fig.save(sprintf(fig_name,"minesOpen"))}
```

## Production capacity over time

```{r}
max_cap_openMines <- df_results %>% 
  left_join(deposit, by="d") %>% mutate(max_prod=mine_open*max_prod_rate) %>% 
  filter(t<2051) %>% group_by(t) %>% reframe(max_prod=sum(max_prod))

df_results %>% 
  filter(t<2051) %>% 
  ggplot(aes(t,cap_total))+
  geom_area(aes(fill=d,group=d))+
  geom_line(data=max_cap_openMines,aes(y=max_prod),linewidth=1,col="red")+
  coord_cartesian(expand = F)+
  labs(x="",y="Production \n Capacity \n [ktons]",
       caption="Red line: Maximum production capacity of opened mines")+
  theme(legend.position = "none",
        axis.text.x = element_text(hjust=1))
if (save_figures){f.fig.save(sprintf(fig_name,"prodCapacity"))}
```

## Curtailed demand or not met

```{r}
slack %>% 
  filter(t<2051) %>% 
  ggplot(aes(t,value))+
  geom_line(linewidth=1)+
  coord_cartesian(expand = F)+
  labs(x="",title="Global demand NOT met, or slack [ktons]",y="")+
  theme(legend.position = "none",
        axis.text.x = element_text(hjust=1))
if (save_figures){f.fig.save(sprintf(fig_name,"demandNotMet"))}
```


# Shadow prices

Some interesting results showcasing shadow prices (or duals). They represent the marginal change in the objective function (total costs) that a marginal change in the restriccion does.

## Demand constraint

Should be positive: more demand implies more cost.

```{r}
# get max with slack non zero
max_slack <- sp_demand %>% filter(value<bigM_cost*0.9) %>% 
  pull(value) %>% max()

# Remove effect of discount rate
discounter <- tibble(t=2022:(t_size+2021),r=(1+discount_rate)^(0:(t_size-1)))

sp_demand %>% 
  filter(t<2051) %>% 
  left_join(discounter) %>% 
  mutate(value=value*r) %>%
  ggplot(aes(t,value))+
  geom_line()+
  # ylim(0,max_slack*1.1)+
  labs(x="",y="Shadow Price \n Demand",caption="All costs in 2022USD (no discount).")
if (save_figures){f.fig.save(sprintf(fig_name,"SP_demand"))}

# note: all should be negative to make sense
# sp_rest %>% reframe(sp_maxProdRate=max(sp_maxProdRate),
#                     sp_rampUp=max(sp_rampUp),
#                     sp_capacity=max(sp_capacity))
```

## Reserve constraint

Should be zero (not binding) or negative

First part:
```{r}
sp_reserve %>% 
  filter(sp_reserve!=0) %>% # remove zeros
  left_join(deposit,by="d") %>% 
  ggplot(aes(reorder(Deposit_Name,sp_reserve),sp_reserve,fill=Resource_Type))+
  geom_col()+
  coord_flip()+
  scale_fill_manual(values = resource_colors)+
  # scale_x_discrete(breaks = NULL)+
  labs(x="Deposit",y="Shadow Price Reserves",fill="")+
  
  theme(axis.text.y = element_text(size=6))
if (save_figures){f.fig.save(sprintf(fig_name,"SP_reserve"))}
```

Second part of reserves (more expensive)

```{r}
cat("Mines that reached depletion: ")
print(sp_reserve %>% filter(sp_reserve2<0) %>% nrow())

sp_reserve %>% 
  filter(sp_reserve2!=0) %>% # remove zeros
  left_join(deposit,by="d") %>% 
  ggplot(aes(reorder(Deposit_Name,sp_reserve2),sp_reserve2,fill=Resource_Type))+
  geom_col()+
  coord_flip()+
  # scale_x_discrete(breaks = NULL)+
  scale_fill_manual(values = resource_colors)+
  labs(x="Deposit",y="Shadow Price Reserves",fill="")+
  theme(axis.text.y = element_text(size=6))
if (save_figures){f.fig.save(sprintf(fig_name,"SP_reserve2"))}
```


## Max Current Production Constraint

Should be negative

```{r}
# ORDER IS KEY TO GET CORRECT SHADOW PRICES FROM THE MATRIX/VECTOR
p1 <- sp_rest %>% 
  filter(t<2051) %>% 
  left_join(deposit,by="d") %>% 
  left_join(dplyr::select(df_results,d,t,mine_open),by=c("d","t")) %>% 
  # filter(mine_open==1) %>% 
  mutate(sp_maxProdRate=mine_open*sp_maxProdRate) %>%
  ggplot(aes(t,Deposit_Name,fill=sp_maxProdRate))+
  geom_tile()+
  # facet_grid(Resource_Type~.,space = "free",scales = "free")+
  ggforce::facet_col(facets = vars(Resource_Type), 
                     scales = "free_y", 
                     space = "free") +
  coord_cartesian(expand = F)+
  scale_fill_gradientn(colors = c("red", "white"))+
  # scale_y_continuous(breaks = NULL)+
  labs(x="",y="Deposit",fill="Shadow Price \nMax prod. rate")+
  theme(axis.text.y = element_text(size=6))
p1+aes(fill=sp_capacity)+labs(fill="Shadow Price \nCapacity")
if (save_figures){f.fig.save(sprintf(fig_name,"SP_Capacity"))}
```

## Max Production Rate Constraint

Should be negative

```{r}
cat("Mines that reached maximum allowable production capacity: ")
# consider only open mines
print(sp_rest %>% 
  filter(t<2051) %>% 
    left_join(dplyr::select(df_results,d,t,mine_open),by=c("d","t")) %>%
    mutate(sp_maxProdRate=mine_open*sp_maxProdRate) %>% 
    filter(sp_maxProdRate < 0) %>% 
   pull(d) %>% unique() %>% length())

# df_results %>% 
#   left_join(deposit, by="d") %>%
#   filter(abs(cap_total-max_prod_rate)<0.1) %>% 
#   pull(d) %>% unique() %>% length()
p1
if (save_figures){f.fig.save(sprintf(fig_name,"SP_MaxProdRate"))}
```


## Max Ramp Up Production Rate Constraint

Should be negative

```{r}
cat("Mines that reached ramp up constraint at some point: ")
print(sp_rest %>% 
  filter(t<2051) %>% filter(sp_rampUp < 0) %>% 
   pull(d) %>% unique() %>% length())

p1+aes(fill=sp_rampUp)+labs(fill="Shadow Price \nRamp up")
if (save_figures){f.fig.save(sprintf(fig_name,"SP_RampUp"))}
```

